---
title: "Assignment"
author: "Nguyen Hoang Nhat Linh"
date: "2025-11-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(ggplot2) # this is primarly used to better visualise the visualisations
```

# Load Data

First of all, we begin by reading the datasets from its respective CSV files. Afterwards, we inspect on each datasets to see the total rows and columns. This allow us to know the original dataset characteristics.

```{r}
# Read csv files with the directory well set beforehand
engine <- read.csv("Engine.csv")
auto <- read.csv("Automobile.csv")
maint <- read.csv("Maintenance.csv")

# Inspect 'Engine'' data
print("Engines:")
str(engine)
summary(engine)
```

```{r}
# Inspect 'Automobile' data
print("Automobile:")
str(auto)
summary(auto)
```

```{r}
# Inspect 'Maintenance' data
print("Maintenance:")
str(maint)
summary(maint)
```

Using "str()", the overall structure of each datasets are well displayed.From the summaries above, we can see that the dataset overview can be seen as follow:

- **Engine**: 8 variables with 88 rows
- **Automobile**: 13 variables with 204 rows
- **Maintenance**: 7 variables with 374 rows 

This is the overall structure when run the datasets in R. From here, we would proceed to solve each tasks.


# Task 1:  

## 1.1: Write the code to inspect the data structure and present the data: The missing values in the dataset were written as "?", replace any "?" with NA; Write code to check: after replacing, how many rows were affected in total? Does this change alter the data distribution?

### Count "?" and replace it with "NA"

In R, the question mark **'?'** is not recognize by as a missing value. This could cause errors in statistical analysis. Therefore, by replacing **'?' to 'NA'** we can code to find out the missing values.

In this question, using the *"mutate()"* function with *"(across(where(is.character)))"*,we can scan all character columns and replace all the '?' to 'NA'.

```{r}
# Count "?" and replace it with "NA"
NA_replace <- function(df, name) {
  # Count rows with "?"
  rows_with_q <- df %>% 
    filter(if_any(everything(), ~ . == "?")) %>% 
    # counts total number of rows that contain at least one instance '?' (everything)
    nrow()
  
  # Replace "?" with NA
  df_clean <- df %>%
    mutate(across(where(is.character), ~na_if(., "?"))) 
  # scan all character columns and replace all the '?' to 'NA'
  
  print(paste("Number of rows affected in", name, ":", rows_with_q))
  return(df_clean)
}

# Apply cleaning
engines <- NA_replace(engine, "Engines")
autos <- NA_replace(auto, "Automobiles")
maint <- NA_replace(maint, "Maintenance")
```

In total, there are 6 rows being affected (all in 'Engine').

```{r}
# Total NA
print(paste("Total NAs in Engine:", sum(is.na(engines))))
print(paste("Total NAs in Automobile:", sum(is.na(autos))))
print(paste("Total NAs in Engine:", sum(is.na(maint))))
```

As we can see, there are 6 '?' and when we replace it, we end up with **34 NAs in total**.

Here, when we convert it to NA, we are simply marking the missing spots properly so R can handle them. All the real numbers and categories stays exactly the same, so the overall distribution of the dataset does not change.




















